High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-03-07 inst/manuscript/manuscript.Rmd
Edits to ms and si, plus commit of many si figures

- name: Lyndon Estes
  affilnum: '1'
  email: lestes@clarku.edu

- name: Su Ye
  affilnum: '1,2'

- name: Lei Song
  affilnum: '1'

- name: Boka Luo
  affilnum: '1'

- name: Ron Eastman
  affilnum: '1'

- name: Zhenhua Meng
  affilnum: '1'

- name: Qi Zhang
  affilnum: '1'

- name: Dennis McRitchie
  affilnum: ''

- name: Stephanie Debats
  affilnum: ''

- name: Justus Muhando
  affilnum: '3'

- name: Angeline Hokar
  affilnum: '3'

- name: Charles Juma 
  affilnum: '3'

- name: Jackson Makuru
  affilnum: '3'

- name: Fredrick Matasa
  affilnum: '3'

- name: Ben Mbatia
  affilnum: '3'

- name: Judith Mbaye
  affilnum: '3'

- name: Isaac Muasa
  affilnum: '3'

- name: Julius Mucha
  affilnum: '3'

- name: Adelide Mugami
  affilnum: '3'

- name: Purent Oduor
  affilnum: '3'

- name: Jeff Ochieng
  affilnum: '3'

- name: Francis Wambua
  affilnum: '3'

- name: Thuo Wanjiku
  affilnum: '3'

- name: Joe Wanyoike 
  affilnum: '3'

- name: Ryan Avery
  affilnum: '4'

- name: Kelly Caylor
  affilnum: '4-6'

@@

- affilnum: 1
  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA, USA

@@

- affilnum: 2
  affil: SpatialCollective, Nairobi, Kenya

- affilnum: 3
  affil: Department of Geography, University of California Santa Barbara, Santa Barbara, CA, USA

- affilnum: 4
  affil: Earth Research Institute, University of California Santa Barbara, Santa Barbara, CA, USA

- affilnum: 5
  affil: Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA, USA

@@

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
abstract: Understanding agricultural change requires reliable, frequently updated maps that describe the characteristics of croplands. Such data are often unavailable for regions dominated by smallholder agricultural systems, which are particularly challenging for remote sensing. To overcome these challenges, we designed a system to minimize several sources of error that arise when mapping smallholder croplands. To overcome errors caused by mismatches between image resolution and cropland scales, the system converts daily, 3.7 m PlanetScope imagery into two seasonal composites within a single agricultural year. To reduce errors that occur when training classifiers, we built a labelling platform that rigorously assesses label accuracy, and creates more accurate consensus labels that train a Random Forests model. The labelling platform and model interact within an active learning process that boosts the accuracy of the resulting cropland probability map, which combines with a segmentation process to delineate individual field boundaries. We applied this system to map Ghana’s croplands for the year 2018. We divided Ghana into 16 mapping regions (12,160-23,535 km$^2$), training separate models for each using a total of 6,299 labels, plus 1,600 for validation. Using an independent map reference sample (n=1,207), we found that overall accuracies of the resulting cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%. Croplands covered 16.1-23.2% of the mapped area, comprising 1,131,146 total fields with an average size of 3.92 ha. Unbiased estimates based on the map reference sample yield indicate a cropland percentage of 17.1% (15.4-18.9%) or 17.6% (15.6-19.6%), depending on whether the cropland probability or field boundary map was assessed. Using the labellers’ digitized field boundaries to estimate biases in field boundary statistics, we calculated an adjusted mean field size of 1.73 ha and total field count of 1,662,281. Although the cropland class contained substantial errors, the system was effective in mitigating error and quantifying resulting performance gains. By minimizing training errors, consensus labelling improved the model’s F1 scores by up to 25%, while 3 iterations of active learning increased the F1 score by 9.1%, on average, which was 2.3% higher than training models with randomly selected labels. Map accuracy may be further improved by replacing Random Forests with a convolutional neural network. These results demonstrate a readily adapted, transferrable framework for developing high resolution, annual, nation-scale maps that provide important details about smallholder-dominated croplands.

keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
csl: ecology.csl

@@

\doublespace

@@

# Introduction
Amidst all the challenges posed by global change, a particular concern is how agricultural systems will adapt to meet humanity's growing food demands, and the impacts that transforming and expanding food systems will have on societies, economies, and the environment [@searchingerCreatingSustainableFood2019]. Significant efforts are being made to address the various aspects of this challenge, including work on diagnosing and closing yield gaps [e.g. @LickerMindgaphow2010; @lobellCropYieldGaps2009; @muellerClosingYieldGaps2012], expanding and commercializing production [@morrisAwakeningAfricaSleeping2009], and to understand [@davisTropicalForestLoss2020; @kehoeNatureRiskFuture2017; @RulliFoodappropriationlarge2014] and mitigate [@EstesReconcilingagriculturecarbon2016] agriculture's ecological impacts. Answering many of the questions these efforts seek to address depends on reliable data that describes the location and characteristics of cropland [@FritzMappingglobalcropland2015], and how these are changing over time. Unfortunately, the data that do exist are in many places inaccurate or missing. Existing estimates of how much global cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g. @FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013]. Such errors can propagate in subsequent analyses that rely on cropland data as inputs, resulting in potentially misleading answers [@Esteslargeareaspatiallycontinuous2018]. Beyond cropland distributions, few data are available on cropland characteristics such as field size, which provide important information on agricultural systems  [@lesivEstimatingGlobalDistribution2019]. These inadequacies are due to the fact that cropland data in much of the world are derived from remotely sensed landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021]

<!-- - The problem in general  -->
<!--     - Africa in particular -->
<!--     - Types of analyses that depend on cropland maps -->

Cropland mapping over Africa is difficult for several reasons. The primary reason relates to the characteristics of the contintent's smallholder-dominated croplands, where half of all fields are smaller than 1 ha [@lesivEstimatingGlobalDistribution2019]. This size is small relative to the 30-250 m resolution of the sensors typically used in many landcover mapping efforts [e.g. @ChenGloballandcover2015; @sulla-menasheHierarchicalMappingAnnual2019], which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979]. In the latter case, the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018]. On top of the matter of scale is 1) high intra-class variability of the cropland class, compounded by the fact that these particular croplands can be heavily intergraded with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and 2) the substantial temporal variability within croplands, both within and between seasons. These latter two aspects pose challenges for the classification algorithms that are applied to the imagery.

These problems arising from cropland characteristics are increasingly being overcome due to technological advances. Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017]. This high spatial *and* temporal resolution addresses the sensor-field scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale. These capabilities are aleady being used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].

@@

We describe here a cropland mapping system that follows these requirements on the way to delineating field boundaries. The first requirement is enabled by the recent availability of CubeSat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017]. Although these data, currently collected by Planet's CubeSat fleet, are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country to continental scale image mosaics to be created for multiple periods during the crop growing calendars, and capturing the intra-annual variability can be more important for classifying cropland than spectral depth [@Debatsgeneralizedcomputervision2016]. This daily revisit capacity is also important for developing seasonal composites in cloudy regions, where satellites with longer return intervals may fail. Lastly, although this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g. see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009]. This allows labels to be generated on the same imagery processed by the classifier, thereby addressing one of the two needs related to training data (requirement 3).

The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016], re-engineered to run on high performance, cloud-based computing clusters, and following recommended practices for controlling for and measuring the errors that occur when training machine learning models for remote sensing applications [@ElmesAccountingtrainingdata2020]. The classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020]. The training and machine learning components are combined within an active learning framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011]. Our framework automates this interactive approach to label selection, which is effective in boosting the performance of classification models while reducing the overall number of training samples required to achieve a given level of performance [e.g. @hamrouniLocalGlobalTransfer2021; @DebatsIntegratingactivelearning2017]. Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, resulting in a vectorized field boundary map that provides important information on field geometry.

We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of climates and agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].

@@

## System components
### Image compositing
The image processing components of our system was designed to work with PlanetScope Analytic surface reflectance data [@PlanetTeamPlanetApplicationProgram2017]. PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency. Although these images are already pre-preprocessed and corrected for atmospheric effects, there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016]. To correct for these factors, we developed a procedure for creating temporal composites representing the primary growing and non-growing seasons within a single year.

PlanetScope imagery is accessed via the Planet API, and an initial order is placed for all imagery falling within the mapping geography and the date ranges for the two compositing periods. The imagery is collected and transferred directly to the system's cloud storage (Amazon Web Services [AWS] S3).

Individual images are then transformed into "analysis ready data" (ARD) [@DwyerAnalysisReadyData2018], by cropping each downloaded scene into a 0.05 degree tile, regardless of cloud cover. Tiles are organized within a larger 1 degree resolution grid that covers the entire continent, which defines the minimum mapping area of interest (AOI; Figure \ref{fig:aois}).

@@

```{r aois, echo = FALSE, out.width="100%", fig.cap="The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
knitr::include_graphics('figures/figure2.png')
```
The temporal compositing process is applied to the tiled daily images for the time period of interest, which in this case was one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year. Imagery from two seasons helps to improve the performance of cropland classifiers [@Debatsgeneralizedcomputervision2016], while having the seasons in same year helps to minimize differences caused by land change. For each pixel in each image in each ARD temporal stack for a given season, two weights are calculated:

@@

\begin{equation} \label{eq:shadow}  
\mathrm{W2_t} =\begin{cases}
    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_1}$, $\mathrm{NIR_2}$, ..., $\mathrm{NIR_i}$\}}.\\
    1, & \text{otherwise}.
  \end{cases}
\end{equation}

@@

Which is a weighted mean for each pixel for each band *B* for the particular compositing period. The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview} 1), where they are stored as cloud-optimized geotiffs, and a "slippy map" rendering is created for each composite using Raster Foundry$\footnote{https://github.com/raster-foundry/raster-foundry}$. The web-rendered imagery is presented within the training data platform (next section).

@@

##### Mapping assignments
Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps. Following a set of pre-defined digitizing rules (see SI), the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays. To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field. The labeller assigns each polygon a class category (e.g. annual cropland), and upon completing all fields submits the assignment to the database. In cases where the target grid cell does not contain any fields, the labeller simply submits the assignment to mark it complete. The labeller is then directed to the next available assignment from a different labelling task.

```{r labeller, echo = FALSE, out.width="100%", fig.cap="An overview of mapper's interface", fig.align='center'}
knitr::include_graphics('figures/figure3.png')
```

@@

Where *i* indicates the particular assignment, and $\beta_{0-5}$ represent varying weights that sum to 1. *I* refers to "inside the box" accuracy, *O* is the accuracy of those portions of the labeller's polygons extending beyond the target grid boundaries, *F* is fragmentation accuracy, a measure of how many individual polygons the labeller delineated relative to the reference, *E* measures how closely each polygon's boundary matched its corresponding reference polygon boundary, and *C* assesses the accuracy of the labeller's thematic labels (see SI for individual formulae). Equation \ref{eq:qaqc} is an extension of the approach described by Estes et al. (2016).

@@

### Classification pipeline
Upon completion of a batch of labels, the platform automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens of instances, depending on the size of the AOI.

#### Feature extraction
The first step is the extraction of additional features from each seasonal image composite. Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016]. We followed this logic in this study, but were constrained to use a smaller feature space because the storage and memory requirements for our mapping geographies in this case were several orders of magnitude larger. For this implemention, we thus extract a set of 16 features, which are the 11X11 mean and 5X5 standard deviation of each band for both seasons (initial tests revealed these window sizes most effective), resulting in 24 overall features with the bands (Table 1).

@@

Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`$\footnote{https://github.com/locationtech/geotrellis}$, `rasterio`$\footnote{https://rasterio.readthedocs.io/en/latest/}$, and `RasterFrames`$\footnote{https://www.cogeo.org/}$. These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features, and convert these into `RasterFrames`, a data structure designed for handling raster data within Apache Spark. Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs$\footnote{https://www.cogeo.org/}$ (COGs), which allows rapid reading of image subsets.

@@

After fitting, the model is applied to the features of the model validation set, and a set of performance metrics are calculated, including binary accuracy, the F1 score (the geometric mean of user's and producer's accuracy), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].

>>>>

- name: L.D. Estes
  affilnum: '1'
  email: lestes@clarku.edu

- name: S. Ye
  affilnum: '1,2'

- name: L. Song
  affilnum: '1'

- name: B. Luo
  affilnum: '1,3'

- name: J. Ronald Eastman
  affilnum: '1,3'

- name: Z. Meng
  affilnum: '1'

- name: Q. Zhang
  affilnum: '1'

- name: D. McRitchie
  affilnum: ''

- name: S. Debats
  affilnum: ''

- name: J. Muhando
  affilnum: '4'

- name: A. Hokar
  affilnum: '4'

- name: C. Juma 
  affilnum: '4'

- name: J. Makuru
  affilnum: '4'

- name: F. Matasa
  affilnum: '4'

- name: B. Mbatia
  affilnum: '4'

- name: J. Mbaye
  affilnum: '4'

- name: I. Muasa
  affilnum: '4'

- name: J. Mucha
  affilnum: '4'

- name: A. Mugami
  affilnum: '4'

- name: P. Oduor
  affilnum: '4'

- name: J. Ochieng
  affilnum: '4'

- name: F. Wambua
  affilnum: '4'

- name: T. Wanjiku
  affilnum: '4'

- name: J. Wanyoike 
  affilnum: '4'

- name: R. Avery
  affilnum: '5'

- name: K. Caylor
  affilnum: '5,6,7'

@@

- affilnum: 1
  affil: Graduate School of Geography, Clark University,  Worcester, MA, USA

@@

- affilnum: 3
  affil: Clark Labs, Clark University, Worcester, MA, USA

- affilnum: 4
  affil: SpatialCollective, Nairobi, Kenya

- affilnum: 5
  affil: Department of Geography, University of California Santa Barbara, Santa Barbara, CA, USA

- affilnum: 6
  affil: Earth Research Institute, University of California Santa Barbara, Santa Barbara, CA, USA

- affilnum: 7
  affil: Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA, USA

@@

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
abstract: Understanding agricultural change requires reliable, frequently updated maps that describe the characteristics of croplands. Such data are often unavailable for regions dominated by smallholder agricultural systems, which are particularly challenging for remote sensing. To overcome these challenges, we designed a system to minimize several sources of error that arise when mapping smallholder croplands. To overcome errors caused by mismatches between image resolution and cropland scales, as well as persistent cloud cover, the system converts daily, 3.7 m PlanetScope imagery into two seasonal composites within a single agricultural year. To reduce errors that occur when training classifiers, we built a labelling platform that rigorously assesses label accuracy, and creates more accurate consensus labels that train a Random Forests model. The labelling platform and model interact within an active learning process that boosts the accuracy of the resulting cropland probability map, which is used in a segmentation process to delineate individual field boundaries. We applied this system to map Ghana’s croplands for the year 2018. We divided Ghana into 16 mapping regions (12,160-23,535 km$^2$), training separate models for each using a total of 6,299 labels, plus 1,600 for validation. Using an independent map reference sample (n=1,207), we found that overall accuracies of the resulting cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%. Croplands covered 16.1-23.2% of the mapped area, comprising 1,131,146 total fields with an average size of 3.92 ha. Estimates based on the map reference sample indicate the cropland percentage is 17.1% (15.4-18.9%) or 17.6% (15.6-19.6%), depending on the map used to estimate the standard error. Using the labellers’ digitized field boundaries to estimate biases in field boundary statistics, we calculated an adjusted mean field size of 1.73 ha and total field count of 1,662,281. Although the cropland class contained substantial errors, the system was effective in mitigating error and quantifying resulting performance gains. By minimizing training errors, consensus labelling improved the model’s F1 scores by up to 25%, while 3 iterations of active learning increased the F1 score by 9.1%, on average, which was 2.3% higher than training models with randomly selected labels. Map accuracy can be improved by replacing Random Forests with a convolutional neural network. These results demonstrate a readily adapted, transferrable framework for developing high resolution, annual, nation-scale maps that provide important details about smallholder-dominated croplands.

keywords: active learning, CubeSat, smallholder, cropland, segmentation, Ghana, consensus labels, machine learning

bibliography: 
  - references.bib
csl: ecology.csl

@@

\singlespace

@@

# Introduction
Amidst all the challenges posed by global change, a particular concern is how agricultural systems will adapt to meet humanity's growing food demands, and the impacts that transforming and expanding food systems will have on societies, economies, and the environment [@searchingerCreatingSustainableFood2019]. Significant efforts are being made to address the various aspects of this challenge, including work on diagnosing and closing yield gaps [e.g. @LickerMindgaphow2010; @lobellCropYieldGaps2009; @muellerClosingYieldGaps2012], expanding and commercializing production [@morrisAwakeningAfricaSleeping2009], and to understand [@davisTropicalForestLoss2020; @kehoeNatureRiskFuture2017; @RulliFoodappropriationlarge2014] and mitigate [@EstesReconcilingagriculturecarbon2016] agriculture's ecological impacts. Answering many of the questions these efforts seek to address depends on reliable data that describes the location and characteristics of cropland [@FritzMappingglobalcropland2015], and how these are changing over time. Unfortunately, the data that do exist are in many places inaccurate. Existing estimates of how much global cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g. @FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013]. Such errors can propagate in subsequent analyses that rely on cropland data as inputs, resulting in potentially misleading answers [@Esteslargeareaspatiallycontinuous2018]. Beyond cropland distributions, few data are available on key cropland characteristics such as field size, an important variable needed to estimate yield and other key food security variables [@CarlettoGuesstimatesGPStimatesLand2015], and as an indicator of farm size [@SambergSubnationaldistributionaverage2016;@levinFarmSizeLandscape2006], a critical component of rural livelihoods given increasing population densities and longstanding debates about the relationship between farm size and productivity [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].

These informational inadequacies are due to the fact that cropland data in much of the world are derived from remotely sensed landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021]. Cropland mapping over Africa is difficult for several reasons. The primary reason relates to the characteristics of the continent's smallholder-dominated croplands, where half of all fields are smaller than 1 ha [@lesivEstimatingGlobalDistribution2019]. This size is small relative to the 30-250 m resolution of the sensors typically used in many landcover mapping efforts [e.g. @ChenGloballandcover2015; @sulla-menasheHierarchicalMappingAnnual2019], which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979]. In the latter case, the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018]. On top of the matter of scale is 1) high intra-class variability of the cropland class, compounded by the fact that these particular croplands can be heavily intergraded with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and 2) the substantial temporal variability within croplands, both within and between seasons. These latter two aspects pose challenges for the classification algorithms that are applied to the imagery.

These problems arising from cropland characteristics are increasingly being addressed due to technological advances. Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017]. This high spatial *and* temporal resolution addresses the sensor-field scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019]. On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale. These capabilities are aleady being used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].

@@

We describe here a cropland mapping system that follows these requirements, with an emphasis on delineating field boundaries. The first requirement is enabled by the recent availability of CubeSat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017]. Although these data, currently collected by Planet's CubeSat fleet, are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country to continental scale image mosaics to be created for multiple periods during the crop growing calendars, and capturing the intra-annual variability can be more important for classifying cropland than spectral depth [@Debatsgeneralizedcomputervision2016]. This daily revisit capacity is also important for developing seasonal composites in cloudy regions, where satellites with longer return intervals may fail. Lastly, although this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g. see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009]. This allows labels to be generated on the same imagery processed by the classifier, thereby addressing one of the two needs related to training data (requirement 3).

The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016], re-engineered to run on high performance, cloud-based computing clusters with a simplified feature set, and following recommended practices for controlling for and measuring the errors that occur when training machine learning models for remote sensing applications [@ElmesAccountingtrainingdata2020]. The classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020]. The training and machine learning components are combined within an active learning framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011]. Our framework automates this interactive approach to label selection, which is effective in boosting the performance of classification models while reducing the overall number of training samples required to achieve a given level of performance [e.g. @hamrouniLocalGlobalTransfer2021; @DebatsIntegratingactivelearning2017]. Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, resulting in a vectorized field boundary map that provides important information on field geometry.

We demonstrate this approach to map cropland in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].

@@

## System components
### Image compositing
The image processing components of our system were designed to work with PlanetScope Analytic surface reflectance data [@planetteamPlanetApplicationProgram2018]. PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency. Although these images are already pre-preprocessed and corrected for atmospheric effects, there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016]. To correct for these factors, we developed a procedure for creating temporal composites representing the primary growing and non-growing seasons within a single year.

PlanetScope imagery is accessed via the Planet API [@planetteamPlanetApplicationProgram2018], and an initial order is placed for all imagery falling within the mapping geography and the date ranges for the two compositing periods. The imagery is collected and transferred directly to a cloud storage platform (Amazon Web Services [AWS] S3).

Individual images are then transformed into analysis ready data (ARD) [@DwyerAnalysisReadyData2018], by subsetting each downloaded image into 0.05 degree tiles, regardless of cloud cover. Tiles are organized within a larger 1 degree resolution grid that covers the entire continent, which defines the minimum mapping area of interest (AOI; Figure \ref{fig:aois}).

@@

```{r aois, echo = FALSE, out.width="100%", fig.cap="The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
knitr::include_graphics('figures/figure2.png')
```
The temporal compositing process is applied to the tiled daily images for the time period of interest, which in this case was one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year. Imagery from two seasons helps to improve the performance of cropland classifiers [@Debatsgeneralizedcomputervision2016], while having the seasons in the same year helps to minimize differences caused by land change. For each pixel in each image in each ARD temporal stack for a given season, two weights are calculated:

@@

\begin{equation} \label{eq:shadow}  
\mathrm{W2_t} =\begin{cases}
    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_{t1}}$, $\mathrm{NIR_{t2}}$, ..., $\mathrm{NIR_{ti}}$\}}.\\
    1, & \text{otherwise}.
  \end{cases}
\end{equation}

@@

Which is a weighted mean for each pixel for each band *B* for the particular compositing period. The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview} 1), where they are stored as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy_Map}" rendering is created for each composite using Raster Foundry$\footnote{https://github.com/raster-foundry/raster-foundry}$. The web-rendered imagery is presented within the training data platform (next section).

@@

##### Mapping assignments
Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps. Following a set of pre-defined digitizing rules (see SI), the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays. To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field. The labeller assigns each polygon a class category (e.g. annual cropland), and upon completing all fields submits the assignment to the database. In cases where the target grid cell does not contain any fields, the labeller simply submits the assignment to mark it complete. The labeller is then directed to the next available assignment from a different labelling task.

```{r labeller, echo = FALSE, out.width="100%", fig.cap="An overview of the labelling platform's interface", fig.align='center'}
knitr::include_graphics('figures/figure3.png')
```

@@

Where *i* indicates the particular assignment, and $\beta_{0-4}$ represent varying weights that sum to 1. *I* refers to "inside the box" accuracy, *O* is the accuracy of those portions of the labeller's polygons extending beyond the target grid boundaries, *F* is fragmentation accuracy, a measure of how many individual polygons the labeller delineated relative to the reference, *E* measures how closely each polygon's boundary matched its corresponding reference polygon boundary, and *C* assesses the accuracy of the labeller's thematic labels (see SI for individual formulae). Equation \ref{eq:qaqc} is an extension of the approach described by Estes et al. (2016).

@@

### Classification pipeline
Upon completion of a batch of labels, the platform automatically launches an ephemeral Elastic Map Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf} cluster consisting of tens of instances, depending on the size of the AOI.

#### Feature extraction
The first step is the extraction of additional features from each seasonal image composite. Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016]. We followed this logic in this study, but were constrained to use a smaller feature space because the storage and memory requirements for our mapping geographies in this case were several orders of magnitude larger. For this implemention, we thus extract a set of 16 features, which are the mean and standard deviations calculated within an 11X11 and 5X5 moving window, respectively (initial tests revealed these two window sizes to be most effective), resulting in 24 overall features, including the original bands (Table 1).

@@

Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://www.cogeo.org/}. These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features, and convert these into `RasterFrames`, a data structure designed for handling raster data within Apache Spark. Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/} (COGs), which allows rapid reading of image subsets.

@@

After fitting, the model is applied to the features of the model validation set, and a set of performance metrics is calculated, including binary accuracy, the F1 score (the geometric mean of precision and recall), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
